---
title: "Lecture 03: Loading CSVs and dplyr joins"
author: "Albert Y. Kim"
date: "September 16, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
```



## I Hate Excel

* It's slow, especially on Macs
* I hate Microsoft's business model
* Using it for analysis can lead to some disastruous errors
    + [The Economist](http://www.economist.com/blogs/graphicdetail/2016/09/daily-chart-3?fsrc=scn/fb/te/bl/ed/excelerrorsandsciencepapers) article
    + [Reinhart/Rogoff](http://www.bloomberg.com/news/articles/2013-04-18/faq-reinhart-rogoff-and-the-excel-error-that-changed-history) and the Excel error that
    changed history

**Comma-separated values** files are a "liberated" format for spreadsheet data



## What is a CSV file?

* An Excel `.xlsx` file not only contains the data, but also lots of metadata (data about the data) that we don't need.
* A `.csv` file [(example)](http://www.sqlshack.com/wp-content/uploads/2015/02/word-image35.png)  is just data and no fluff:
    + Rows are separated by lines.
    + Values for a given row (i.e. variables) are separated by commas. Each row has equal number of commas.
    + The first row is typically a header row with the column/variable names

Note `.csv` files are in **tidy** data format (from Lec02).



## Convert to CSV

* Typically in the open source community, data is shared in `.csv` format.
* But if someone sends you a `.xlsx` file, there are two ways to convert it to `.csv` format:
    1. Using Excel (necessitates Microsoft Office)
    1. Using Google Docs (free, but necessitates a Gmail address)



## 1. Convert to CSV with Excel

1. Open `states.xlsx` in Excel
1. From Excel's menu bar -> *File* -> *Save As...* -> *File Format* -> *Windows Comma Separated (.csv)*
1. Add `.csv` to the filename to *Save As*. Your filename should be `states.csv`



## 2. Convert to CSV with Google Sheets

1. While logged-in to your gmail, open your [Google Drive](https://drive.google.com/drive/my-drive)
1. Drop the file `states.xlsx` into the browser and double-click it
1. Click on open in Google Sheets
1. From the Google Sheets menu bar go to *File* -> *Download As* -> *Comma-separated values* -> Save the file to your computer.
1. Rename the file `states.csv`



## Load CSV into RStudio

1. In the RStudio Files panel, navigate to the file and click it, and click "Import Dataset"
1. Copy the commands in the Code Preview box in the bottom right. You can paste this an R Script so that you can automate this procedure in the future.
1. Click *Import*

Note: `readr::read_csv()` is an improved version of `read.csv()` in base (default) R.



## Directory Structure

Let's look at the `read_csv()` command. Note it specifies the **directory
path**. If you don't specify this path, R will be default look for the file in the
**current working directory**, which you can

* Set/identify with `setwd()`/`getwd()`
* Set/identify in the Files panel -> *More* -> Set As/Go To Working Directory

This often trips up people when sharing code.








## Combining Data Sets via Joins

Imagine we have two data frames `x` and `y`:

```{r, echo=FALSE}
x <- data_frame(x1=c("A","B","C"), x2=c(1,2,3))
y <- data_frame(x1=c("A","B","D"), x3=c(TRUE,FALSE,TRUE))
x 
y
```


## Combining Data Sets via Joins

We want to **join** them along the categorical variable `x1` and end up with a
new data frame that has all three variables.



## Combining Data Sets via Joins


```{r, echo=TRUE}
inner_join(x, y, by = "x1")
full_join(x, y, by = "x1")
```



## Combining Data Sets via Joins


* `dplyr`'s operations to join data sets are [inspired](https://github.com/hadley/dplyr/blob/master/vignettes/two-table.Rmd#databases) by SQL (Structured Query
Language), which is used to query large databases.
* If a value is missing during the join, `NA` (R code for missing value) is
inserted.

In our example, both data frames `x` and `y` the had the "join by" variable
named the same: `x1`. Hence:

```
inner_join(x, y, by = "x1")
```



## The "join by" variable


What if the "join by" variable has different names?  Ex:

* name `a` in data frame `x`
* name `b` in data frame `y`

then run:  `inner_join(x, y, by = c("a"="b"))`. 

Note it is

* ` by = c("a"="b")` and not
* `by = c("a"=="b")`

See `by` in the "Argument" section of the help file `?join`.



## Types of Joins

* See `?join` for a list of the different types of joins
* This
[illustration](https://twitter.com/yutannihilation/status/551572539697143808)
succinctly summarizes all of them.






