---
title: "Lecture 03: Loading CSVs and dplyr joins"
author: "Albert Y. Kim"
date: "September 16, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
```




## Preparation for Today

* Open `UCBAdmissions.xlsx` in Excel



## Today's Data

The University of California Berkeley was sued in 1973 for bias against women who had applied for admission to graduate schools. We consider the $n=4526$ people who applied to the 6 largest departments.

```{r, echo=FALSE}
data("UCBAdmissions")
UCBAdmissions %>% as.data.frame() %>% head()
```





## Loading Data

We will explore two methods for loading data:

* Via a Comma Separated Values (CSV) file, from 
    + Excel files
    + Google docs



## CSV Files

* An Excel `.xlsx` file not only contains data, but also lots of [metadata](https://www.google.com/search?q=metadata&oq=metadata&aqs=chrome..69i57j69i65j0l4.1567j0j9&sourceid=chrome&es_sm=119&ie=UTF-8), i.e. data about data, that we don't need.
* A CSV file is just data, where values for a given row are separated by commas. It is the most flexible way to share a table of data.

CSV = No fluff, just stuff.





## Excel Files

To load data in an Excel spreadsheet in RStudio, we need the values to be in **tidy**
format:

* Data are in a table: i.e. perfect rectangular shpe
* Columns are variables
* Rows are observations, with the exception of the first row which are the variable names AKA a *header* row.





## Excel Files

We convert the `UCBAdmissions.xlsx` file to CSV format to eliminate the metadata:

1. From Excel's menu bar -> *File* -> *Save As...* -> *Format* -> *Windows Comma Separated (.csv)*
1. Add `.csv` to the filename to *Save As*. i.e. your filename should be `UCBAdmissions.csv`
1. Click *Continue*



## Excel Files

We load the CSV file into R.

1. From RStudio -> *Environment* Panel -> *Import Dataset* -> *From Text File...* -> Click on `UCBAdmissions.csv`
1. Make sure *Heading* is set to *Yes*. This indicates that the first row is a *header* row.
1. Click *Import*

Note in your console panel R spits out the command to do this automatically. You can copy this line into your R scripts.



## Google Docs

From the Google Docs menu bar (not your browser's menu bar):

* *File* -> *Download As* -> *Comma-separated values* -> Save the file to your computer.
* Repeat the steps from the last slide to load into R.





## Combining Data Sets via Join Operations

Imagine we have two data frames **`x** and **y`**:

```{r, echo=FALSE}
x <- data.frame(x1=c("A","B","C"), x2=c(1,2,3))
y <- data.frame(x1=c("A","B","D"), x3=c(TRUE,FALSE,TRUE))
x
y
```

We want to **join** them along the categorical variable `x1`and end up with a
new data frame that has all three variables:



## Combining Data Sets via Join Operations


```{r, echo=TRUE}
inner_join(x, y, by = "x1")
full_join(x, y, by = "x1")
```



## Combining Data Sets via Join Operations


* `dplyr`'s operations to join data sets are inspired by SQL (Structured Query
Language), which is used to query large databases.
* If a value is missing during the join, `NA` (R code for missing value) is
inserted.

In our example, both data frames `x` and `y` the had the "join by" variable
named the same: `x1`. Hence:

```
inner_join(x, y, by = "x1")
```



## The "join by" variable


If the "join by" variable has different names in the data frames.  Ex:

* name `a` in data frame `x`
* name `b` in data frame `y`

then run:  `inner_join(x, y, by = c("a"="b"))`. 

Note it is

* ` by = c("a"="b")` and not
* `by = c("a"=="b")`

See `by` in the "Argument" section of the help file `?join`.



## Types of Joins

* See `?join` for a list of the different types of joins
* This
[illustration](https://twitter.com/yutannihilation/status/551572539697143808)
succinctly summarizes all of them.



## Good Practice


At the beginning of `Lec03.R`, I load all necessary packages at the beginning.

This is good practice people know which packages they need 
installed. For example, I didn't know I needed a particular package before I 
boarded a 4 hour flight because the package loading was buried in
the file.



