---
title: "Lecture 10: Categorical Variables with forcats"
author: "Albert Y. Kim"
date: "October 5, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
```




## Categorical variables in R

Categorical variables coded as `factor`s in R. They consists of:

* **data**: a vector of values
* **metadata**: `levels` that determine an ordering of the categories

In particular, the `levels` determine the order that categorical variables
are shown in plots.




## Categorical variables in R

For the longest time, the fact that the `levels` were metadata i.e. data about the data
that isn't visible in "tidy data", made working with `factor`s a real pain in R.

Until August 31, 2016 upon the release of the `forcats` package.




## Example

Let's look at today's data, a sample of categorical variables from the General Social survey.
Run the following in R

```{r, eval=FALSE}
library(dplyr)
library(forcats)
data("gss_cat")
glimpse(gss_cat)
```




## Example

In particular, let's look at `relig` i.e. religion of respondent. Run:

```{r, eval=FALSE}
summary(gss_cat$relig)
levels(gss_cat$relig)
```

The `levels()` function shows the default ordering.




## Working with Factors

There are two chief modifications people want to do to factors:

1. Change the order of the levels for visualizations via `fct_reorder()`
1. Change the values of a factor via `fct_recode()`




## Resources

* In R, type `fct_` and scroll the auto-complete to see all functions
* Intro to `forcats` on [RStudio blog](https://blog.rstudio.org/2016/08/31/forcats-0-1-0/)
* More in depth treatment in Wickham's [R for Data Science](http://r4ds.had.co.nz/factors.html#modifying-factor-levels)








