---
title: "Lecture 18: Leaflet"
author: "Albert Y. Kim"
date: "November 4, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
```



## Leaflet

Today we will create **interactive** maps. Run the following in the console:

```{r, eval=FALSE}
library(leaflet)
leaflet() %>%
  addTiles() %>%
  addMarkers(lng=-73.1755, lat=44.01005, popup="Warner Hall")
```



## Open Street Maps

[OpenStreetMap (OSM)](https://www.openstreetmap.org/#map=5/51.500/-0.100) is a
collaborative project to create a free editable map of the world. This is
crowd-sourced open source alternative to Google Maps. 

For example, the following institutions use it instead of Google Maps:

* Craigslist
* Apple
* Foursquare



## Leaflet

[Leaflet](http://leafletjs.com/) is the leading open-source JavaScript library
for mobile-friendly interactive maps. This exists independent of R.

**Quote on their page**: *Leaflet doesn't try to do everything for everyone. Instead
it focuses on making the basic things work perfectly*.



## Leaflet for R

There is an R package to interface with Leaflet which simply hooks Leaflet into
R: <https://rstudio.github.io/leaflet/>



## Basic Usage

You create a Leaflet map with these basic steps:

> 1. Create a map widget by calling `leaflet()`.
> 1. Add layers (i.e. features) to the map by using layer functions (e.g.
`addTiles`, `addMarkers`, `addPolygons`) to modify the map widget. Kind of like the
way you add layers to a `ggplot()`
> 1. Print the map widget to display it.





## Map Widget

If you understand `ggplot()`, you can understand `leaflet()`:

* It sets the base layer, to which we add layers.
* You can specify `data`, which can be
    + a regular data frame with columns
        + `lat` or `latitude`
        + `lng`, `long`, `lng`, or `longitude`
    + a `maps` package data frame
    + an `sp` package `SpatialPolygons` object. See this [link](https://rstudio.github.io/leaflet/shapes.html)







## Other leaflet Features

* Can embed in R Markdown documents
* Can [integrate with Shiny](https://rstudio.github.io/leaflet/shiny.html)





