---
title: "Lecture 05: More ggplot2"
author: "Albert Y. Kim"
date: "September 23, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```




## Today

In `Lec05.R` we look at

* `geom_line()`. In particular the `group` aesthetic
* `geom_smooth()`: a type of `geom_line()`
* `geom_histogram()`




## What is a smoother?

Say you have a sequence of points that appear "noisy":

```{r, echo=FALSE}
simple <- data_frame(
    x = seq(from=0, to=10, by=0.1),
    y = sin(x) + rnorm(length(x), mean = 0, sd = 0.5)
  )
ggplot(data=simple, aes(x=x, y=y)) +
  geom_point()
```






## What is a smoother?

A **smoother** attempts to smooth out the **noise** and pick out the **signal**:

```{r, echo=FALSE}
ggplot(data=simple, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(se=FALSE, col="red")
```



## Linear Regression as a Smoother

**Note**: linear regression is also a form of smoothing! Although in this case not so good:

```{r, echo=FALSE}
ggplot(data=simple, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method="lm",se=FALSE, col="red")
```





## What is a smoother?

Smoothers work by doing a "local regression" in a "sliding window":
[GIF](http://simplystatistics.org/wp-content/uploads/2014/02/loess.gif)

This is called a [lowess](https://en.wikipedia.org/wiki/Local_regression)
(Locally Weighted Scatterplot Smoothing) curve, also called loess curve.



## Today's Data: babynames Package

The [`babynames`](https://github.com/hadley/babynames) package contains for each
year from 1880 to 2013, the number of children born of each sex given each name
in the United States (source: SSA).

Note: only names with more than 5 occurrences in a given year for a given sex are included.










