---
title: "Lecture 06: Finishing ggplot2"
author: "Albert Y. Kim"
date: "September 26, 2016"
output: ioslides_presentation
---

<style>
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
slides > slide.backdrop {
  background: white;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```




## Examples for Today

In `Lec06.R` we study new `geom_`'s:

* `geom_histogram`
* `geom_boxplot()`
* `geom_bar()`. In particular:
    + Understanding `stat=identity` vs `stat=bin`
    + and `position` adjustments



## Data: UC Berkeley Admissions

In 1973, the UC Berkeley was sued for bias against women who had applied for
admission to graduate schools.  n=4526 applicants:

```{r, echo=FALSE}
data(UCBAdmissions)
UCBAdmissions <- UCBAdmissions %>% 
  as.data.frame() %>% 
  group_by(Admit, Gender) %>% 
  summarise(Freq = sum(Freq))
UCBAdmissions
```



## Data: UC Berkeley Admissions

```{r, echo=FALSE}
UCBPlot <- ggplot(data=UCBAdmissions, aes(x=Gender, y=Freq, fill=Admit)) +
  ggtitle("Acceptance Rate Split by Gender") +
  xlab("Gender")
UCBPlot + geom_bar(stat = "identity") +
  ylab("# of Applicants")
```



## Data: UC Berkeley Admissions

```{r, echo=FALSE}
UCBPlot + geom_bar(stat = "identity", position="fill") +
  ylab("Prop of Applicants")
```



## Data: UC Berkeley Admissions

However, there was another variable researchers could consider: `Dept` applied to.

```{r, echo=FALSE}
data(UCBAdmissions)
UCBAdmissions <- as.data.frame(UCBAdmissions) %>% tbl_df()
UCBAdmissions
```



## Data: Today's Exercise

Investigate:

1. How male vs female acceptance varied by department.
2. The "competitiveness" of different departments as measured by acceptance rate.
