---
title: "Babynames"
author: "Albert Y. Kim"
date: "October 3, 2016"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(babynames)
library(ggplot2)
library(dplyr)
```


## Creating App

While there is no "right" way to create such an app, you **definitely** want to
break down the task into bite-sized pieces. For example I

1. Got the ggplot working using "pretend" inputs, **then tested**
1. Made the `name`, `start_year`, and `end_year` input working one-by-one (but didn't connect to plot yet), **then tested**
1. Then synced `input$name`, `input$start_year`, and `input$end_year` into ggplot one-by-one, **then tested**


```{r eruptions, echo=FALSE}
inputPanel(
  # Input name
  textInput("name", label = "Enter Name:", value="Jane"),
  # Input start year
  numericInput("start_year", label = "Start Year:", value = 1880, min = 1880, max = 2014, step = 1),
  # Input end year
  numericInput("end_year", label = "End Year:", value = 2014, min = 1880, max = 2014, step = 1)
)

renderPlot({
  # Pretend Inputs
  input_name <- "Albert"
  input_start_year <- 1880
  input_end_year <- 2014
  
  # Generate Plots
  names <- babynames %>%
    filter(name == input$name & year >= input$start_year & year <= input$end_year )
  ggplot(data=names, aes(x=year, y=prop, col=sex)) + 
    geom_line()
})
```


